Formalizing Stack Safety as a Security Property

### 2021-09-30 submission4/main.tex
Cut WBCF down more

The key intuition is that, to be considered as having returned to {\tt main}, {\tt f} needs
to do more than just execute a return-labeled instruction: it must actually reach the
appropriate return target in {\tt main}.
To capture this intuition, we add a new piece of context to our diagram:
a stack of return targets.  Now a callee is
only considered to have returned when it reaches a state that matches the caller's target.
A return target is associated with a set of sealed locations that will become unsealed
when it is reached, represented here by coloring both the target and its sealed data with the
same color (blue).

In the example, {\tt f} reaches its return target on its first return, to instruction 4.
But {\tt f}'s second return is to the wrong instruction
(4 rather than 6) because the stack pointer has been overwritten (shown in light red).
Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
violates the integrity property.

\paragraph*{Well-bracketed control flow.} Together,
integrity and confidentiality protect a caller's data until its callee
returns to it. However, it is possible that the caller is not
returned {\em to}, but rather returned {\em past}, to some ancestor caller!
For example, function $A$ calls function $B$, which in turn calls $C$, which
returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
and branching to the address after the call from $A$ to $B$.  Strictly
speaking, this is not an information-flow violation, since neither $A$'s nor
$B$'s stack frames are improperly read or written by $C$.  But it is still
arguably unsafe if $A$'s programmer depends on $B$ completing its execution before
it returns.
%
Thus, another aspect of safe use of the stack is that calls
and returns should be correctly nested---e.g., $C$ should
only return to $B$.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
call \emph{well-bracketed control flow (WBCF)}, and is orthogonal to
confidentiality and integrity, so we treat is as an optional additional property.

@@

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs, including RISCy ones.  In
particular, there may be no single ``call'' instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

@@

In other words, stepwise stack confidentiality states that if, at the start of the call,
we vary all stack locations that are supposed to be secret from the callee
(which in our simple setting is the entirety of the stack),
the resulting trace remains in lockstep with the original trace
until either both return together or one failstops. This implies that
if the callee reads a secret location, it must failstop.
%
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the elements in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

Finally, we formalize a version of WBCF, which captures the same informal idea
as that of Skorstengaard et al.\cite{SkorstengaardSTKJFP} using our context machinery.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).

@@

Tags in the PIPE system are not accessible to application code, being separated
from their associated values at the hardware level. So they are not modeled
as part of the values, and therefore our properties make no claims about
their behavior.

>>>>

The key intuition is that, to be considered as having returned to {\tt main}, {\tt f} needs
to do more than just execute a return-labeled instruction: it must actually reach the
appropriate return target in {\tt main}.
To capture this intuition, we add a new piece of context:
a stack of return targets.  Now a callee is
only considered to have returned when it reaches a state that matches the caller's target.
A return target is associated with a set of sealed locations that will become unsealed
when it is reached, represented here by coloring both the target and its sealed data with the
same color (blue).

In the example, {\tt f} reaches its return target on its first return, to instruction 4.
But {\tt f}'s second return is to 4 rather than 6
because the stack pointer has been overwritten (shown in light red).
Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
violates the integrity property.

\paragraph*{Well-bracketed control flow} Together,
integrity and confidentiality protect a caller's data until its callee
returns to it. However, it is possible that the caller is not
returned {\em to}, but rather returned {\em past}, to some ancestor caller!
Strictly speaking, this is not an information-flow violation. But it is still
arguably unsafe if the ancestor depends on the skipped function completing its execution before
it returns.
%
Thus, another aspect of safe use of the stack is that calls
and returns should be correctly nested---a caller should
only return to its own callee.  This is what Skorstengaard et al.\cite{SkorstengaardSTKJFP}
call \emph{well-bracketed control flow (WBCF)}, and is orthogonal to
confidentiality and integrity, so we treat is as an optional additional property.

@@

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs.  There may be no single ``call''
instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

@@

In short, if we consider a variant of the stack at the start of a call, 
the resulting trace remains in lockstep with the original trace
until either both return together or one failstops.
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the elements in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately.

Finally, we formalize WBCF as a simple property of the behavior of returns.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).

@@

Tags in the PIPE system are not accessible to application code, being separated
from their associated values at the hardware level. So they are not modeled
as values, and therefore our properties make no claims about
their behavior.
